\texttt{Taxonomic supertree construction with}\texttt{\emph{ Incertae sedis}}\texttt{ taxa}

### 2018-04-26 doc/incertae_sedis.tex
Rename subsection.
Add interdigitation figure and discuss in naming section.

\subsubsection{Split-based semantics}
\paragraph{Equivalence of a taxonomy tree and a list of rooted splits}

@@

@@

However, in the \emph{incertae sedis} framework, it is possible for one name to
apply to multiple nodes. For example, in Figure \ref{fig:One-name-can}b, the
name $A$ can apply to nodes $x$ and $y$. Here the name $A$ corresponds to the split
$A1\,A2\,|\bullet B\,C$, leaving out $D$ since it is \emph{incertae sedis}. The two nodes $x$
and $y$ display the splits $A1\,A2\,|\bullet B\,C\,D$ and $A1\,A2\,D|\bullet
B\,C$ respectively, and both
of these splits display the split $A1\,A2\,|\bullet B\,C$, so the name $A$ can apply to both
$x$ and $y$. This cannot happen without\emph{ incertae sedis} taxa except at
monotypic nodes. When faced with a choice about where to place a name, our
solution is to find the most tip-ward node where the name can apply and attach
the name to this node. This corresponds to a conservative choice about whether
\emph{incertae sedis} taxa should be added to an existing taxon.
\begin{figure}
\hfill{}\subfloat[Taxonomy tree]{\includegraphics[scale=0.5]{Figures/name1nodes2/tax-shaded}}
\hfill{}\subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/name1nodes2/solution-shaded}}
\hfill{}
\caption{\label{fig:One-name-can}One name can be consistent with multiple nodes.}
\end{figure}

It is also possible for multiple names to apply to a single node. For example,
in Figure \ref{fig:Multiple-names-can}a, the taxon $A$ corresponds to the split
$B1\,B2\,C\,|\bullet\,Y$, and its child taxon $B$ corresponds to the split $B1\,B2\,|\bullet\,Y$. The
edge leading to $A$ is consistent with the split for $B$, but the name $B$ is
applied to the node with the smallest include group. However, in the
solution tree (Fig. \ref{fig:Multiple-names-can}b), there is is only node for
both names $A$ and $B$ to apply to. In this case, we solve this problem by
introducing a monotypic parent, and applying $A$ to the newly created parent node
(Fig. \ref{fig:Multiple-names-can}c).
\begin{figure}
\hfill{}\subfloat[Taxonomy tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/tax-shaded}}
\hfill{}\subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/synth-shaded}}
\hfill{}\subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/solution2-shaded}}
\hfill{}
\caption{\label{fig:Multiple-names-can}Multiple names can apply to a single node.}
\end{figure}

\bdrcomment{Ben: discuss naming + interdigitation}

@@

\bdrcomment{Ben: Add figure}

>>>>

\subsubsection{Terminology for rooted splits}
\paragraph{Equivalence of a taxonomy tree and a list of rooted splits}

@@

\begin{figure}
\hfill{}\subfloat[Taxonomy tree]{\includegraphics[scale=0.5]{Figures/name1nodes2/tax-shaded}}
\hfill{}\subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/name1nodes2/solution-shaded}}
\hfill{}
\caption{\label{fig:One-name-can}One name can be consistent with multiple nodes.}
\end{figure}

@@

However, in the \emph{incertae sedis} framework, it is possible for one name to
apply to multiple nodes. For example, in Figure \ref{fig:One-name-can}b, the
name $A$ can apply to nodes $x$ and $y$. Here the name $A$ corresponds to the split
$A1\,A2\,|\bullet B\,C$, leaving out $D$ since it is \emph{incertae sedis}. The two nodes $x$
and $y$ display the splits $A1\,A2\,|\bullet B\,C\,D$ and $A1\,A2\,D|\bullet
B\,C$ respectively, and both
of these splits display the split $A1\,A2\,|\bullet B\,C$, so the name $A$ can apply to both
$x$ and $y$. This cannot happen without\emph{ incertae sedis} taxa except at
monotypic nodes. When faced with a choice about where to place a name, our
solution is to find the most tip-ward node where the name can apply and attach
the name to this node. This corresponds to a conservative choice about whether
\emph{incertae sedis} taxa should be added to an existing taxon.

\begin{figure}
\hfill{}\subfloat[Taxonomy tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/tax-shaded}}
\hfill{}\subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/synth-shaded}}
\hfill{}\subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/solution2-shaded}}
\hfill{}
\caption{\label{fig:Multiple-names-can}Multiple names can apply to a single node.}
\end{figure}

It is also possible for multiple names to apply to a single node. For example,
in Figure \ref{fig:Multiple-names-can}a, the taxon $A$ corresponds to the split
$B1\,B2\,C\,|\bullet\,Y$, and its child taxon $B$ corresponds to the split $B1\,B2\,|\bullet\,Y$. The
edge leading to $A$ is consistent with the split for $B$, but the name $B$ is
applied to the node with the smallest include group. However, in the
solution tree (Fig. \ref{fig:Multiple-names-can}b), there is is only node for
both names $A$ and $B$ to apply to.  This kind of situation arises when a taxon
becomes monotypic after its \emph{incertae sedis} children are placed elsewhere.
For such cases, we can solve this problem by (i) selecting a taxon at the
node that is ancestral to all other taxa and (ii) moving this taxon to a
newly-introduced out-degree-1 parent.  For example, $A$ is moved to an
out-degree-1 parent in Fig. \ref{fig:Multiple-names-can}c.  We repeatedly perform
this procedure until no taxon at the node is the parent of all taxa.

However, there are cases where thus procedure does not result in one name per
node.  This case can occur, for example, when two incertae sedis siblings are 
``inter-digitated'', or  combined into a single clade in such a way that their
members cannot be separated.  When the members of the two clades share the same
MRCA, as in Fig. \ref{fig:interdigitation}, then the names for both clades
apply to the same node.  However, in this case, neither name is ancestral to
the other, and the previous solution does not apply.  In such cases, we choose
one of the names arbitrarily.

\begin{figure}
\hfill{}\subfloat[Taxonomy tree]{\includegraphics[scale=0.5]{Figures/is-interdigitation/tax-shaded}}
\hfill{}\subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/is-interdigitation/synth-shaded}}
\hfill{}
\caption{\label{fig:interdigitation}Interdigitation of \emph{incertae sedis} siblings.  Here the names $A$ and $B$ apply to the same node on the solution tree, but neither is ancestral to the other on the taxonomy.}
\end{figure}

@@

\bdrcomment{Ben: discuss figure \ref{fig:interdigitation}.}
